---
title: '**MixCluster**'
output: html_document
---


*Mixture Model of Gaussian copulas to cluster mixed data*

**Description:**
  
* *Authors*: **Matthieu Marbac** and **Christophe Biernacki** and **Vincent Vandewalle**.
* *License*: [GPL-2](http://www.gnu.org/licenses/gpl-2.0.html).
*  *Download MixClust 1.0*: [link](https://r-forge.r-project.org/R/?group_id=1939).
* *Reference*: [Model-based clustering of Gaussian copulas for mixed data](http://arxiv.org/abs/1405.1299), Marbac, M. and Biernacki, B. and Vandewalle, V., 2015, preprint.

<a id="top"></a>
**Site map:**  

* <a href="#intro">Introduction</a>.
* <a href="#tutorial">Tutorial (Example 2.2 of the preprint)</a>.

<a id="intro"></a>

### Introduction
*VarSelLCM* achieves the mixed data clustering with a Gaussian copula mixture model, since copulas, and in particular the Gaussian ones, are powerful tools for easily modelling the distribution of multivariate variables. 
Indeed, considering a mixing of continuous, integer and ordinal variables (thus all having a cumulative distribution function), this mixture model allows to model intra-component dependencies like a Gaussian mixture, so with standard correlation meaning.
Simultaneously, it preserves standard margins associated to continuous, integer and ordered features, namely the Gaussian, the Poisson and the ordered multinomial distributions.
As an interesting by-product, the proposed mixture model generalizes many well-known ones and provides tools of visualization based on the model parameters.
At a practical level, the Bayesian inference is retained and it is achieved with a Metropolis-within-Gibbs sampler.

Tool functions (*summary* and *MixClusVisu*) facilitate the result interpretation.

* <a href="#top">Go to the top</a>


<a id="tutorial"></a>

###  Overview of the VarSelLCM functions

This section reproduces Example 2.2 of the preprint 

First, the cluster analysis is performed with two components, then a model summary is given (partition and parameters). Finally, the cluster interpretation can be done based on the graphical and numerical presentations of the parameters.

**Loadings**

```{r, comment=""}
rm(list=ls())
set.seed(135)
library(MixCluster)

# Loading of a dataset simulated from a bi-component mixture model of Gaussian copulas
# (see Example 2.2 page 6)
# The first column indicates the class membership
# The last three column are used for the clustering
data(simu)
```

**Clustering of mixed variables**

```{r, comment=""}
# Cluster analysis by the bi-component mixture model of Gaussian copulas
# without constrain between the correlation matrices
res.mixclus <- MixClusClustering(simu[,-1], 2)
```

**Confusion matrix**

```{r, comment=""}
# Confusion matrix between the estimated (row) and the true (column) partition
table(res.mixclus@data@partition, simu[,1])
```


**Model Summary**

```{r, comment=""}
# Summary of the model
summary(res.mixclus)
```

**Visualisation**
We reproduce the elements of Figure 1
```{r, comment=""}
# Visualisation
# Update of the results (computing the conditional expectations of the latent vectors
# related to the Gaussian copulas)
res.mixclus <- MixClusUpdateForVisu(res.mixclus)
```



```{r, comment=""}
# Scatterplot of the individuals  (Figure 1.(a)) described by three variables:
# one continuous (abscissa), one integer (ordiate) and one binary (symbol).
# Colors indicate the component memberships
plot(simu[,2:3], col=simu[,1], pch=16+simu[,4], xlab=expression(x^1), ylab=expression(x^2))
```


```{r, comment=""}
# Scatterplot of the individuals in the first PCA-map of the first-component of the model
MixClusVisu(res.mixclus, class = 1, figure = "scatter", xlim=c(-10,4), ylim=c(-4,4))
```


```{r, comment=""}
# Correlation circle of the first PCA-map of the first-component of the model
MixClusVisu(res.mixclus, class = 1, figure = "circle")
```

* <a href="#top">Go to the top</a>